CFF GCT Study

### 2025-01-17 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
removed FEV1pp calculations

    "Highest FEV1" = Highest.FEV1..last.12.months..FEV1..L.,
    "Second Highest FEV1" =
      Second.Highest.FEV1..12.months...FEV1..2nd.highest.in.L..,

@@

    `Highest FEV1`, `Second Highest FEV1`,

@@

# Calculate FEV1 pp and height, weight, and BMI percentiles
lung_function <- df %>%
    ID, Age, Gender, Race, `Height (cm)`, `Highest FEV1`,
    `Second Highest FEV1`
  filter(!is.na(Gender))
lung_function$`Average of Highest FEV1s` <-
  rowMeans(lung_function[, c("Highest FEV1", "Second Highest FEV1")], na.rm = T)
lung_function$Gender <- factor(lung_function$Gender,
  levels = c("Male", "Female"),
  labels = c("male", "female")

@@

lung_function$Race <- factor(lung_function$Race,
  levels = c(
    "White", "Black/African American",
    "Asian Non-Indian", "Asian Indian", "American Indian/Alaska Native",
    "Native Hawaiian/Other Pacific Islander"
  ),

@@

    "Caucasian", "African-American", "NE Asian", "SE Asian", "Other/mixed",
    "Other/mixed"

@@

lung_function$`Highest FEV1pp (2012)` <- pctpred_GLI(
  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
  gender = lung_function$Gender, ethnicity = lung_function$Race,
  FEV1 = lung_function$`Highest FEV1`

@@

lung_function$`Second Highest FEV1pp (2012)` <- pctpred_GLI(
  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
  gender = lung_function$Gender, ethnicity = lung_function$Race,
  FEV1 = lung_function$`Second Highest FEV1`
lung_function$`Average of Highest FEV1pp (2012)` <- pctpred_GLI(
  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
  gender = lung_function$Gender, ethnicity = lung_function$Race,
  FEV1 = lung_function$`Average of Highest FEV1s`

lung_function$`Highest FEV1pp (2022)` <- pctpred_GLIgl(
  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
  gender = lung_function$Gender, FEV1 = lung_function$`Highest FEV1`
)
lung_function$`Second Highest FEV1pp (2022)` <- pctpred_GLIgl(
  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
  gender = lung_function$Gender, FEV1 = lung_function$`Second Highest FEV1`
)
lung_function$`Average of Highest FEV1pp (2012)` <- pctpred_GLIgl(
  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
  gender = lung_function$Gender,
  FEV1 = lung_function$`Average of Highest FEV1s`
)

df <- left_join(
  df,
  lung_function %>%
    select(ID, `Highest FEV1pp (2012)`:`Second Highest FEV1pp (2022)`)
)
# Height, weight, BMI percentiles
df$`Height %ile` <- sds(df$`Height (cm)`,
  age = ifelse(df$Age > 20, 20, df$Age),
  item = "height2_20", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",
  sex = as.character(df$Gender)
)
df$`Weight %ile` <- sds(df$`Weight (kg)`,
  age = ifelse(df$Age > 20, 20, df$Age),
  item = "weight2_20", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",
  sex = as.character(df$Gender)
)
df$`BMI %ile` <- sds(df$BMI,
  age = ifelse(df$Age > 20, 20, df$Age),
  item = "bmi", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",
  sex = as.character(df$Gender)
)
df <- df %>% select(-`Highest FEV1`, -`Second Highest FEV1`)
# Time from OGTT to GCT
df$OGTT.Date.Time <- mdy_hm(df$OGTT.Date.Time)
df$Study.visit.PFTs.Date. <- mdy_hm(df$Study.visit.PFTs.Date.)
df$`Days From OGTT to GCT` <-
  as.numeric(as.Date(df$OGTT.Date.Time) - as.Date(df$Study.visit.PFTs.Date.))
df$OGTT.Date.Time <- NULL
df$Study.visit.PFTs.Date. <- NULL
# Per Christine, we can exclude C046's 140 value and keep them in the NGT
# category because they had multiple normal OGTTs after this one and the 140 is
# likely a transcription error
df$`OGTT 0 min`[df$ID == "C046"] <- NA
# Get diagnosis
df$Diagnosis <- "NGT"
df$Diagnosis[df$`OGTT 0 min` >= 100] <- "IFG"
df$Diagnosis[df$`OGTT 60 min` >= 200] <- "INDET"
df$Diagnosis[df$`OGTT 120 min` >= 140] <- "IGT"
df$Diagnosis[df$`OGTT 120 min` >= 200] <- "CFRD"
df[which(rowSums(is.na(df[, grep("OGTT", colnames(df))])) == 3), "Diagnosis"] <-
  NA
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD")
)
df$Dysglycemia <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "AGT")
)
df$Diabetes <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
  labels = c("NGT", "Pre-CFRD", "Pre-CFRD", "Pre-CFRD", "CFRD")
)
df$CFRD <- factor(df$`OGTT 120 min` >= 200,
  levels = c(F, T),
  labels = c("No", "Yes")
)
# Has OGTT vs. does not
df$`Got OGTT and GCT?` <- apply(
  df[, c("OGTT 120 min", "GCT 60 min plasma")], 1, function(r) {
    !any(is.na(r))
  }
)
df$`Got OGTT and GCT?` <- factor(df$`Got OGTT and GCT?`,
  levels = c(T, F),
  labels = c(
    "Have GCT 1 Hour Plasma or OGTT 2 Hour",
    "Missing GCT 1 Hour Plasma or OGTT 2 Hour"
  )
)
# Write for Christine to check
write.csv(df,
  file = "./Christine Chan/CFF GCT Study/Data_Clean/gct_analysis_data.csv",
  row.names = F, na = ""
)
```

>>>>

    "Highest FEV1pp" = Highest.FEV1..last.12.months....FEV1.highest....,
    "Second Highest FEV1pp" =
      Second.Highest.FEV1..12.months....FEV1.2nd.highest....,
    "Average of Highest FEV1pp" =
      Average.of.the.two.highest.FEV1.s...Average.FEV1.......,

@@

    `Highest FEV1pp`, `Second Highest FEV1pp`, `Average of Highest FEV1pp`,

@@

# Height, weight, BMI percentiles
df$`Height %ile` <- sds(df$`Height (cm)`,
  age = ifelse(df$Age > 20, 20, df$Age),
  item = "height2_20", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",
  sex = as.character(df$Gender)

@@

df$`Weight %ile` <- sds(df$`Weight (kg)`,
  age = ifelse(df$Age > 20, 20, df$Age),
  item = "weight2_20", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",
  sex = as.character(df$Gender)
df$`BMI %ile` <- sds(df$BMI,
  age = ifelse(df$Age > 20, 20, df$Age),
  item = "bmi", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",
  sex = as.character(df$Gender)
# Time from OGTT to GCT
df$OGTT.Date.Time <- mdy_hm(df$OGTT.Date.Time)
df$Study.visit.PFTs.Date. <- mdy_hm(df$Study.visit.PFTs.Date.)
df$`Days From OGTT to GCT` <-
  as.numeric(as.Date(df$OGTT.Date.Time) - as.Date(df$Study.visit.PFTs.Date.))
df$OGTT.Date.Time <- NULL
df$Study.visit.PFTs.Date. <- NULL
# Per Christine, we can exclude C046's 140 value and keep them in the NGT
# category because they had multiple normal OGTTs after this one and the 140 is
# likely a transcription error
df$`OGTT 0 min`[df$ID == "C046"] <- NA
# Get diagnosis
df$Diagnosis <- "NGT"
df$Diagnosis[df$`OGTT 0 min` >= 100] <- "IFG"
df$Diagnosis[df$`OGTT 60 min` >= 200] <- "INDET"
df$Diagnosis[df$`OGTT 120 min` >= 140] <- "IGT"
df$Diagnosis[df$`OGTT 120 min` >= 200] <- "CFRD"
df[which(rowSums(is.na(df[, grep("OGTT", colnames(df))])) == 3), "Diagnosis"] <-
  NA
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD")
df$Dysglycemia <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "AGT")
df$Diabetes <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
  labels = c("NGT", "Pre-CFRD", "Pre-CFRD", "Pre-CFRD", "CFRD")
df$CFRD <- factor(df$`OGTT 120 min` >= 200,
  levels = c(F, T),
  labels = c("No", "Yes")
# Has OGTT vs. does not
df$`Got OGTT and GCT?` <- apply(
  df[, c("OGTT 120 min", "GCT 60 min plasma")], 1, function(r) {
    !any(is.na(r))
df$`Got OGTT and GCT?` <- factor(df$`Got OGTT and GCT?`,
  levels = c(T, F),

@@

    "Have GCT 1 Hour Plasma or OGTT 2 Hour",
    "Missing GCT 1 Hour Plasma or OGTT 2 Hour"

@@

# Write for Christine to check
write.csv(df,
  file = "./Christine Chan/CFF GCT Study/Data_Clean/gct_analysis_data.csv",
  row.names = F, na = ""

@@

```
